% CREATED BY DAVID FRISK, 2014
\begin{thebibliography}{69}
%\bibitem{Template} Frisk, D. (2014) \textit{A Chalmers University of Technology Master's thesis template for} \LaTeX . Unpublished.

% NOTE:
% References are not automatically sorted in order of appearance, so make sure to insert new references at the correct spot!

% Introduction
\bibitem{swift} Chris Lattner and Apple Inc. (2014) \textit{Apple Swift Language} [online] Available at: \url{https://developer.apple.com/swift/} (retrieved 2015-05-12)

\bibitem{rust} Graydon Hoare and Rust Project Developers. (2010) \textit{Rust Language} [online] Available at: \url{http://www.rust-lang.org} (retrieved 2015-05-12)

\bibitem{haskell} haskell.org. (1990) \textit{Haskell Language} [online] Available at: \url{https://www.haskell.org/} (retrieved 2015-05-13)

\bibitem{erlang} Joe Armstrong. (1986) \textit{Erlang Language} [online] Available at: \url{http://www.erlang.org/} (retrieved 2015-05-13)

% Background
\bibitem{ninenines} Joe Armstrong. (2007) \textit{What's all this fuss about Erlang?} [online] Available at: \url{https://pragprog.com/articles/erlang} (retrieved 2015-05-12)
\bibitem{elixir}  José Valim. (2012) \textit{Elixir} [online] Available at: \url{http://elixir-lang.org} (retrieved 2015-05-12)
\bibitem{lfe}  Robert Virding. (2008) \textit{Lisp Flavored Erlang} [online] Available at: \url{http://lfe.io} (retrieved 2015-05-12)
\bibitem{cloudhaskell}  Jeff Epstein. (2011) \textit{Cloud Haskell} [online] Available at: \url{http://haskell-distributed.github.io} (retrieved 2015-05-12)
\bibitem{yhc}  Dimitry Golubovsky. (2008) \textit{York Haskell Compiler: Haskell on BEAMs} [online] Available at: \url{http://yhc06.blogspot.se/2008/05/haskell-on-beams.html} (retrieved 2015-05-12)
\bibitem{haskerl}  Torbjörn Törnkvist. (2015) \textit{Haskerl: Haskell subset compiled to the Beam machine} [online] Available at: \url{https://github.com/etnt/Haskerl} (retrieved 2015-06-01)

% Background parser
\bibitem{bnfc} Markus Forsberg, Aarne Ranta. (2005) \textit{The Labelled BNF Grammar Formalism} [online] Available at: \url{http://bnfc.digitalgrammars.com/LBNF-report.pdf} (retrieved 2015-05-09)

% Background code gen?
\bibitem{CoreErlangIntro} Carlsson, R. (2001) \textit{An introduction to Core Erlang} [online] Available at: \url{http://www.it.uu.se/research/group/hipe/cerl/doc/cerl_intro.ps} (retrieved 2015-03-16)


% Method
\bibitem{Scrum} Scrum.Org and ScrumInc (2014) \textit{The Scrum Guide} [online] Available at: \url{http://www.scrumguides.org/scrum-guide.html} (retrieved 2015-03-16)
\bibitem{git}  Linus Torvalds. (2015) \textit{Git} [online] Available at: \url{http://git-scm.com} (retrieved 2015-05-12)
\bibitem{alex} Chris Dornan, Isaac Jones, and Simon Marlow. (2003) \textit{Alex User Guide} [online] Available at: \url{https://www.haskell.org/alex/doc/alex.pdf} (retrieved 2015-05-09)
\bibitem{happy} Simon Marlow and Andy Gill. (2001) \textit{Happy User Guide} [online] Available at: \url{https://www.haskell.org/happy/doc/happy.pdf} (retrieved 2015-05-09)
\bibitem{lex} Lesk, Michael E., and Eric Schmidt. (1975) \textit{Lex: A lexical analyzer generator}.
\bibitem{yacc} Johnson, Stephen C. (1975) \textit{Yacc: Yet another compiler-compiler}. Murray Hill, NJ. Bell Laboratories.

% Impl grammar
% Moved to method

% Impl depcheck?
\bibitem{interfacefiles}  GHC Compiler Wiki. (2015) \textit{Interface files} [online] Available at: \url{https://ghc.haskell.org/trac/ghc/wiki/Commentary/Compiler/IfaceFiles} (retrieved 2015-05-12)

% Impl type checker
\bibitem{TypeSchemes} Damas, L. and Milner, R. (1982) \textit{Principal type-schemes for functional programs} [online] Available at: \url{http://web.cs.wpi.edu/~cs4536/c12/milner-damas_principal_types.pdf} (retrieved 2015-02-05)

\bibitem{FunTutorial} Peyton Jones, S. and Lester, D. (1992) \textit{Implementing functional languages: a tutorial} [online] Available at: \url{http://research.microsoft.com/en-us/um/people/simonpj/Papers/pj-lester-book/} (retrieved 2015-05-05)


% Impl code gen
\bibitem{CoreErlang} Pérez, D. C. and Garzía, H. F. (2008) \textit{The CoreErlang package} [online] Available at: \url{https://hackage.haskell.org/package/CoreErlang} (retrieved 2015-03-16)

% Disc tc
\bibitem{implfun} Peyton Jones, S. (1987) \textit{The Implementation of Functional Programming Languages} [online] Available at: \url{http://research.microsoft.com/en-us/um/people/simonpj/papers/slpj-book-1987/start.htm} (retrieved 2015-03-16)

\end{thebibliography}
