\section{Parsing source code} \label{sec:bnfc}

\todo{Assigned David}

% A programming language is similar to a natural language in that it (is accompanied by)/has/(is described by) a grammar. Programming languages are however different in that they are built from their grammar and are thus unambiguous. // I think this says almost the same thing but focuses less on the spoken/natural language? use parts or whole if you'd like

Similar to a spoken language a programming language is built from a grammar, but different from a spoken language a programming language's grammar is exact. This comes from that a spoken language's grammar is a construction in retrospect and also that it does not always follow the rules. 

\todo{HELP! Grammar, words} 

\todo{Is this a correct statement?}

\todo{mention that programming language grammar is (more) strict and exact} 

A programming language's grammar needs to be exact for it to not allow syntax to be ambiguous. The grammar describes ways in which the smaller pieces of the language can be put together to form bigger ones, often with more meaning. From this grammar a lexer is produced. The lexers job is to match all words in the source code to tokens in the grammar. For example in english the word 'boat' would match a noun token.

\todo{maybe try to rearrange/rephrase the passage above to give it more of a "röd tråd" or better flow, will try to help with this later }

The next step is to combine all these tokens to expressions or sentences. This is done with a parser. To make the language as expressive as possible you build up bigger expressions out of smaller expressions. This will create a tree structure that will be easier to work with. 

To save a lot of work the lexer and parser can be automatically generated from a tool when the grammar is written in Backus-Naur-form or BNF. This BNF converter, or BNFC for short \todo{maybe glossary}, is developed at Chalmers and used in earlier courses meaning that the group members was already famililar with the tool \todo{plt ref here}.

\todo{Add appendix with example code and its representation}

