\chapter{Design and Implementation}

\todo{Assigned Johan L}

This chapter describes the design and implementation of the Hopper language and its compiler with all the compiler components. There will be overviews of each part describing the choices made and the reasoning behind those choices.

\todo{Brief dscription of the grammar, features, maybe misc topics of interest that dont have any immediate section to be written in}

When first designing Hopper there was already an outline of what the language should aim to be. The type safety and clean syntax of Haskell combined with the powerful framework for concurrency and parallelism supplied by the Erlang virtual machine was the key features from which Hopper would be built. Within those confines there were further choices as to what features to include, and not to include, to make the language relevant and useful.

\todo{Need help figuring out what features that are relevant to mention here!}

Features.
Interoperability with Erlang
General-purpose (we can assume this?)
Strict/Eager instead of lazy
Type checking (already talked about enough?)
Pure?
ADTs
Module stuff?


Discuss the most relevant features here...

Grammar.

\todo{Talk about grammar here or in D&I: Parser?}

\todo{delete D&I: the hopper language}

%\input{include/Designandimpl_hopper}
\input{include/Designandimpl_pipeline}
\input{include/Designandimpl_parser}
\input{include/Designandimpl_renamer}
\input{include/Designandimpl_depcheck}
\input{include/Designandimpl_typechecker}
\input{include/Designandimpl_codegen}
