\section{Grammar}

Using the \gls{bnfc} to generate a lexer and parser from our grammar was overall very successful. BNFC has all the features wanted for implementing the product. It halved the work by producing both a lexer and a parser which would otherwise have had to be written separately from scratch. It also has indentation sensitivity built in which is used to insert extra tokens to match against. The group had some problems at the beginning of the project to understand when this rule activated, but they were solved by studying the generated source code. There is a simple way to test if a problem in the grammar is due to either the layout rule or a grammatical rule. This is done by overriding the layout rule by inserting the tokens manually to check if the specified grammar still behaves correctly.

At the start of the project there were discussions in the group on whether or not to use BNFC to generate the lexer and parser. Choosing to get an easy start with the generated lexer and parser was accompanied by the notion that the group might want to switch to using Alex \cite{alex} and Happy \cite{happy} directly at a later stage. This did not come to happen during the project but might still be something to consider for the future. Implementing the group's own lexer and parser would mean greater control over the process.

The rewrite done half way through the project temporarily held up the introduction of new features. It was a necessary change however; new syntax could not otherwise have been added without ambiguity.
